---
title: "Aides Drug"
author: "Cyndhell Anne Pilarca"
date: "July 24, 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(dplyr)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aids drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$: The vaccine (drug) is not effective in preventing aids among drug users.   
$H_A$: The vaccine (drug) is effective in preventing aids among drug users.

Data for this study came from a clinical trial for the drug triUmpsal.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

The response variable analysis that I'll be using is cat~cat because because the aids and the vaccine have the subgroup got both categorical. Because the question is about drug users, we have to subset the data to include only druggies.

Graphical display: bar chart, numerical display, the inferential, kai square 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->


```{r}
DruggieData = filter(triUmp, DrugUser == "yes")
```


## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~infected + treatment,data=DruggieData)
barchartGC(~infected + treatment,data=DruggieData, type="percent")

```

The graph shows there is a different between the groups with aids and the group that none, among the people who did got the treatment have no different with those who got the placebo effect.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data= DruggieData)
rowPerc(table1)
colPerc(table1)
```

Among all those people who are infected in aids those who did not get the vaccine are 3 times more likely to get the aids with the placebo than the people who got the treatment. The people who have aids have a 24.29% chances to get treated than the people who got the placebo.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value 0.0000001242 from chi square tests shows that the vaccine is effective, p-value is to small that's why we reject the null hypothesis. The fisher exact test value which is odds is 0.35 times less likely to get aids than someone who gets the placebo.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Is the vaccine able to help people with aids and those people who don't receive the treatment. In the conclusion that we gather on the research the effectiveness of the vaccine in preventing aids have higher treatment that placebo effect. 


